---
title             : "ELIKTU analysis of outliers (younger cohort)"
shorttitle        : "outliers"
author:
  - name          : "Denis Matrov"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "jazzhole@gmail.com"

affiliation:
  - id            : "1"
    institution   : "Tallinn University"
    
keywords          : ["outliers", "eliktu"]
wordcount         : "X"

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, echo = FALSE, include = FALSE, cache=FALSE}
library("papaja")
library("ggplot2")
library("scales")
library("broom")
library("kableExtra")
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(
  cache = TRUE,
  cache.comments = FALSE,
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  echo = FALSE,
  strip.white = TRUE,
  comment = "#>",
  fig.path = "../figures/",
  results = "asis",
  tidy = "styler",
  dev = c('pdf', 'tiff'),
  dev.args = list(pdf = list(colormodel = 'cmyk'), 
                  tiff = list(compression = 'lzw')),
  dpi = 600,
  fig.width = 7,
  fig.height = 7,
  concordance = TRUE,
  global.par = TRUE
)

par(font.main = 1, cex.main = 1.05)
options(knitr.table.format = "latex")
theme_set(theme_apa(base_size = 15) + theme(legend.position = "bottom"))
```

```{r multiv_function}
multv_outliers <- function(data, print_n = FALSE, tolBAC=0.000000001,
                           toladj=0.05, tolDDC=0.0001) {
  chunk <- na.omit(data)
  O3d <- O3prep(chunk[, -1], method=c("BAC", "adjOut", "DDC"),
                tolBAC=tolBAC, toladj=toladj, tolDDC=tolDDC)
  O3d1 <- O3plotM(O3d, caseNames = chunk$kood)
  if(print_n) {print(O3d1$nOut)}
  grid.arrange(O3d1$gO3 + theme(plot.margin = unit(c(0, 1, 0, 0), "cm")), O3d1$gpcp, ncol=1, heights=c(2,1))
}
```


# Multivariaate analysis of outliers (younger cohort)

```{r data-load}
library(here)
library(tidyverse)
library(psych)
library(ggthemes)
library(gridExtra)
library(OutliersO3)
options(scipen=999)
path <- here("mod_data", "AMIS_toitumine_wide.csv")
data <- read_csv(path)[, c(1:3, 125:312)]
data <- subset(data, kohort==1, -c(2, 3))
data <- data[, -grep("_9", colnames(data))]
data <- data[, -grep("_33", colnames(data))]
data <- cbind(data[, 1], data[, -1][, order(colnames(data[-1]))])
n <- length(colnames(data))
```


## Vitamins E and K
see Table \@ref(tab:vitamins-EK-table)

```{r vitam_EK}
chunk <- data[, c(1, (n-3):n)]
```

(ref:vitaminsEK) Multivariate outlier analysis for data on vitamins E and K. 

```{r vitamins-EK, results='asis', fig.cap="(ref:vitaminsEK)", fig.height=11, fig.width=10, cache=TRUE}
multv_outliers(chunk)
```


```{r vitamins-EK-table, cache=TRUE}
knitr::kable(subset(data, kood %in% c(1447), c(1, (n-3):n)), row.names = FALSE, caption = "One case with significantly outlying scores for vitamins E and K.")
```

## Vitamin D
see Table \@ref(tab:vitamins-D-table)

```{r vitam_D}
chunk <- data[, c(1, (n-7):(n-5))]
```

(ref:vitaminD) Multivariate outlier analysis for data on vitamin D. 

```{r vitamins-D, results='asis', fig.cap="(ref:vitaminD)", fig.height=11, fig.width=10, cache=TRUE}
multv_outliers(chunk)
```


```{r vitamins-D-table, cache=TRUE}
knitr::kable(subset(data, kood %in% c(1317, 2035, 1469), c(1, (n-7):(n-5))), row.names = FALSE, caption = "Three cases with significantly outlying scores for vitamin D.")
```

## Vitamin C
see Table \@ref(tab:vitamins-C-table)

```{r vitam_C}
chunk <- data[, c(1, (n-10):(n-8))]
```

(ref:vitaminsC) Multivariate outlier analysis for data on vitamin C. 

```{r vitamins-C, results='asis', fig.cap="(ref:vitaminsC)", fig.height=11, fig.width=10, cache=TRUE}
multv_outliers(chunk)
```

```{r vitamins-C-table, cache=TRUE}
knitr::kable(subset(data, kood %in% c(1440, 1719), c(1, (n-10):(n-8))), row.names = FALSE, caption = "Two cases with significantly outlying scores for vitamin C.")
```


## Vitamin B6
see Table \@ref(tab:vitamins-B6-table)

```{r vitam_B6}
chunk <- data[, c(1, (n-13):(n-11))]
```

(ref:vitaminsB6) Multivariate outlier analysis for data on vitamin B6. 

```{r vitamins-B6, results='asis', fig.cap="(ref:vitaminsB6)", fig.height=11, fig.width=10, cache=TRUE}
multv_outliers(chunk)
```

```{r vitamins-B6-table, cache=TRUE}
knitr::kable(subset(data, kood %in% c(1466), c(1, (n-13):(n-11))), row.names = FALSE, caption = "One case with significantly outlying scores for vitamin B6.")
```

## Vitamin B2
see Table \@ref(tab:vitamins-B2-table)

```{r vitam_B2}
chunk <- data[, c(1, (n-16):(n-14))]
```

(ref:vitaminsB2) Multivariate outlier analysis for data on vitamin B2. 

```{r vitamins-B2, results='asis', fig.cap="(ref:vitaminsB2)", fig.height=11, fig.width=10, cache=TRUE}
multv_outliers(chunk)
```


```{r vitamins-B2-table, cache=TRUE}
knitr::kable(subset(data, kood %in% c(1466), c(1, (n-16):(n-14))), row.names = FALSE, caption = "One case with significantly outlying scores for vitamin B2.")
```

## Vitamin B12
see Table \@ref(tab:vitamins-B12-table)

```{r vitam_B12}
chunk <- data[, c(1, (n-19):(n-17))]
```

(ref:vitaminsB12) Multivariate outlier analysis for data on vitamin B12. 

```{r vitamins-B12, results='asis', fig.cap="(ref:vitaminsB12)", fig.height=11, fig.width=10, cache=TRUE}
multv_outliers(chunk)
```

```{r vitamins-B12-table, cache=TRUE}
knitr::kable(subset(data, kood %in% c(1524, 1926, 1122, 1904, 2083), c(1, (n-19):(n-17))), row.names = FALSE, caption = "Five cases with significantly outlying scores for vitamin B12.")
```

## Vitamin B1
see Table \@ref(tab:vitamins-B1-table)

```{r vitam_B1}
chunk <- data[, c(1, (n-22):(n-20))]
```

(ref:vitaminsB1) Multivariate outlier analysis for data on vitamin B1. 

```{r vitamins-B1, results='asis', fig.cap="(ref:vitaminsB1)", fig.height=11, fig.width=10, cache=TRUE}
multv_outliers(chunk)
```

```{r vitamins-B1-table, cache=TRUE}
knitr::kable(subset(data, kood %in% c(1466), c(1, (n-22):(n-20))), row.names = FALSE, caption = "One case with significantly outlying scores for vitamin B1.")
```

## Vitamin A
Data for the 18 years of age is missing entirely.
See Table \@ref(tab:vitamins-A-table)

```{r vitam_A}
chunk <- data[, c(1, (n-25), (n-23))]
```

(ref:vitaminsA) Multivariate outlier analysis for data on vitamin A. 

```{r vitamins-A, results='asis', fig.cap="(ref:vitaminsA)", fig.height=11, fig.width=10, cache=TRUE}
multv_outliers(chunk)
```

```{r vitamins-A-table, cache=TRUE}
knitr::kable(subset(data, kood %in% c(1904, 2083, 1986, 2065), c(1, (n-25), (n-23))), row.names = FALSE, caption = "Four cases with significantly outlying scores for vitamin A.")
```

## Vegetables
Vegetables data has no cases of putative outliers.

## Sweets
see Table \@ref(tab:sweets-table)

```{r sweets}
chunk <- data[, c(1, (n-33):(n-31))]
```

(ref:sweets-fig) Multivariate outlier analysis for data on sweets. 

```{r sweets-figure, results='asis', fig.cap="(ref:sweets-fig)", fig.height=11, fig.width=10, cache=TRUE}
multv_outliers(chunk)
```

```{r sweets-table, cache=TRUE}
knitr::kable(subset(data, kood %in% c(1352), c(1, (n-33):(n-31))), row.names = FALSE, caption = "One case with significantly outlying scores for sweets.")
```

## Retinol and salt
see Table \@ref(tab:retinol-table)

```{r retinol}
chunk <- data[, c(1, (n-37), (n-36), (n-34))]
```

(ref:retinol-fig) Multivariate outlier analysis for data on retinol and salt. 

```{r retinol-figure, results='asis', fig.cap="(ref:retinol-fig)", fig.height=11, fig.width=10, cache=TRUE}
multv_outliers(chunk)
```

```{r retinol-table, cache=TRUE}
knitr::kable(subset(data, kood %in% c(1122, 2083), c(1, (n-37), (n-36), (n-34))), row.names = FALSE, caption = "Two cases with significantly outlying scores for retinol and salt.")
```

## Proteins

Protein data has no cases of putative outliers.

## Poultry and Pantothen
No data for younger cohort at 15 years of age for poultry or pantothen. 
Otherwise the multivariate data looks OK.

## Other drinks
see Table \@ref(tab:drinks-table)

```{r drinks}
chunk <- data[, c(1, (n-50):(n-48))]
```

(ref:drinks-fig) Multivariate outlier analysis for data on non-alcoholic drinks. 

```{r drinks-figure, results='asis', fig.cap="(ref:drinks-fig)", fig.height=11, fig.width=10, cache=TRUE}
multv_outliers(chunk)
```

```{r drinks-table, cache=TRUE}
knitr::kable(subset(data, kood %in% c(1475, 1733), c(1, (n-50):(n-48))), row.names = FALSE, caption = "Two cases with significantly outlying scores for non-alcoholic drinks.")
```

## Niacin

`Niacineq_18` has no values for the younger cohort. 
If I take the variable `Niatsiini_eqv_18` instead, there are no obvious outliers.

## Dairy

Dairy data has no cases of putative outliers.

## Meat

Meat data has no cases of putative outliers.

## Lipids

Lipids data has no cases of putative outliers.

## LDL
see Table \@ref(tab:ldl-table)

```{r ldl}
chunk <- data[, c(1, (n-72):(n-70))]
```

(ref:ldl-fig) Multivariate outlier analysis for data on LDL. 

```{r ldl-figure, results='asis', fig.cap="(ref:ldl-fig)", fig.height=11, fig.width=10, cache=TRUE}
multv_outliers(chunk)
```

```{r ldl-table, cache=TRUE}
knitr::kable(subset(data, kood %in% c(1793), c(1, (n-72):(n-70))), row.names = FALSE, caption = "One cases with significantly outlying scores for LDL.")
```

## HOMA
see Table \@ref(tab:homa-table)

```{r homa}
chunk <- data[, c(1, (n-75):(n-73))]
```

(ref:homa-fig) Multivariate outlier analysis for data on HOMA index. 

```{r homa-figure, results='asis', fig.cap="(ref:homa-fig)", fig.height=11, fig.width=10, cache=TRUE}
multv_outliers(chunk)
```


```{r homa-table, cache=TRUE}
knitr::kable(subset(data, kood %in% c(2106, 1065, 1279, 1227), c(1, (n-75):(n-73))), row.names = FALSE, caption = "Four cases with significantly outlying scores for HOMA index.")
```

## HDL

HDL data has no cases of putative outliers.

## Fruits and berries
see Table \@ref(tab:fruit-table)

```{r fruit}
chunk <- data[, c(1, (n-81):(n-79))]
```

(ref:fruit-fig) Multivariate outlier analysis for data on fruits and berries. 

```{r fruit-figure, results='asis', fig.cap="(ref:fruit-fig)", fig.height=11, fig.width=10, cache=TRUE}
multv_outliers(chunk)
```


```{r fruit-table, cache=TRUE}
knitr::kable(subset(data, kood %in% c(2099), c(1, (n-81):(n-79))), row.names = FALSE, caption = "One case with significantly outlying scores for fruits and berries.")
```

## Folate

Folate data has no cases of putative outliers.

## Fish

Fish data has a lot of 0 values, the multivariate test fails to converge.

## Fats

Fats data has no cases of putative outliers.

## Eggs
see Table \@ref(tab:eggs-table)

```{r eggs}
chunk <- data[, c(1, (n-93):(n-91))]
```

(ref:eggs-fig) Multivariate outlier analysis for data on eggs. 

```{r eggs-figure, results='asis', fig.cap="(ref:eggs-fig)", fig.height=11, fig.width=10, cache=TRUE}
multv_outliers(chunk)
```


```{r eggs-table, cache=TRUE}
knitr::kable(subset(data, kood %in% c(1279, 1802), c(1, (n-93):(n-91))), row.names = FALSE, caption = "Two cases with significantly outlying scores for eggs.")
```


## CHL
see Table \@ref(tab:chl-table)

```{r chl}
chunk <- data[, c(1, (n-100):(n-98))]
```

(ref:chl-fig) Multivariate outlier analysis for data on CHL. 

```{r chl-figure, results='asis', fig.cap="(ref:chl-fig)", fig.height=11, fig.width=10, cache=TRUE}
multv_outliers(chunk)
```


```{r chl-table, cache=TRUE}
knitr::kable(subset(data, kood %in% c(1880, 1385, 1793), c(1, (n-100):(n-98))), row.names = FALSE, caption = "Three cases with significantly outlying scores for CHL.")
```

## Cereal

Cereal data has no cases of putative outliers.

## Carbohydrates

Carbohydrates data has no cases of putative outliers.

## Biotin
see Table \@ref(tab:biotin-table)

```{r biotin}
chunk <- data[, c(1, (n-113):(n-112))]
```

(ref:biotin-fig) Multivariate outlier analysis for data on biotin. 

```{r biotin-figure, results='asis', fig.cap="(ref:biotin-fig)", fig.height=11, fig.width=10, cache=TRUE}
multv_outliers(chunk)
```


```{r biotin-table, cache=TRUE}
knitr::kable(subset(data, kood %in% c(1226, 1802), c(1, (n-113):(n-112))), row.names = FALSE, caption = "Two cases with significantly outlying scores for biotin.")
```

## Alcohol

Alcohol data has a lot of 0 values, the multivariate test fails to converge.
