---
title             : "ELIKTU analysis of outliers (older cohort)"
shorttitle        : "outliers"
author:
  - name          : "Denis Matrov"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "jazzhole@gmail.com"

affiliation:
  - id            : "1"
    institution   : "Tallinn University"
    
keywords          : ["outliers", "eliktu"]
wordcount         : "X"

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, echo = FALSE, include = FALSE, cache=FALSE}
library("papaja")
library("ggplot2")
library("scales")
library("broom")
library("kableExtra")
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(
  cache = TRUE,
  cache.comments = FALSE,
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  echo = FALSE,
  strip.white = TRUE,
  comment = "#>",
  fig.path = "../figures/",
  results = "asis",
  tidy = "styler",
  dev = c('pdf', 'tiff'),
  dev.args = list(pdf = list(colormodel = 'cmyk'), 
                  tiff = list(compression = 'lzw')),
  dpi = 600,
  fig.width = 7,
  fig.height = 7,
  concordance = TRUE,
  global.par = TRUE
)

par(font.main = 1, cex.main = 1.05)
options(knitr.table.format = "latex")
theme_set(theme_apa(base_size = 15) + theme(legend.position = "bottom"))
```

```{r multiv_function}
multv_outliers <- function(data, print_n = FALSE, tolBAC=0.000000001,
                           toladj=0.05, tolDDC=0.0001) {
  chunk <- na.omit(data)
  O3d <- O3prep(chunk[, -1], method=c("BAC", "adjOut", "DDC"),
                tolBAC=tolBAC, toladj=toladj, tolDDC=tolDDC)
  O3d1 <- O3plotM(O3d, caseNames = chunk$kood)
  if(print_n) {print(O3d1$nOut)}
  grid.arrange(O3d1$gO3 + theme(plot.margin = unit(c(0, 1, 0, 0), "cm")), O3d1$gpcp, ncol=1, heights=c(2,1))
}
```


# Multivariaate analysis of outliers (Older cohort)

```{r data-load}
library(here)
library(tidyverse)
library(psych)
library(ggthemes)
library(gridExtra)
library(OutliersO3)
options(scipen=999)
path <- here("mod_data", "AMIS_toitumine_wide.csv")
data <- read_csv(path)[, c(1, 125:312)]
data <- data[, -grep("_9", colnames(data))]
data <- cbind(data[, 1], data[, -1][, order(colnames(data[-1]))])
n <- length(colnames(data))
```


## Vitamins E and K
$~$

```{r vitam_EK}
chunk <- data[, c(1, (n-5):n)]
```

(ref:vitaminsEK) Multivariate outlier analysis for data on vitamins E and K. 

```{r vitamins-EK, results='asis', fig.cap="(ref:vitaminsEK)", fig.height=11, fig.width=10, cache=TRUE}
multv_outliers(chunk)
```


```{r vitamins-EK-table, cache=TRUE}
knitr::kable(subset(data, kood %in% c(1899, 1019), c(1, (n-5):n)), row.names = FALSE, caption = "Two cases with significantly outlying scores for vitamins E and K.")
```

## Vitamin C
$~$

```{r vitam_C}
chunk <- data[, c(1, (n-14):(n-11))]
```

(ref:vitaminsC) Multivariate outlier analysis for data on vitamin C. 

```{r vitamins-C, results='asis', fig.cap="(ref:vitaminsC)", fig.height=11, fig.width=10, cache=TRUE}
multv_outliers(chunk)
```


```{r vitamins-C-table, cache=TRUE}
knitr::kable(subset(data, kood %in% c(1995), c(1, (n-14):(n-11))), row.names = FALSE, caption = "One case with significantly outlying scores for vitamin C.")
```


## Vitamin B6
$~$

```{r vitam_B6}
chunk <- data[, c(1, (n-18):(n-15))]
```

(ref:vitaminsB6) Multivariate outlier analysis for data on vitamin B6. 

```{r vitamins-B6, results='asis', fig.cap="(ref:vitaminsB6)", fig.height=11, fig.width=10, cache=TRUE}
multv_outliers(chunk)
```


```{r vitamins-B6-table, cache=TRUE}
knitr::kable(subset(data, kood %in% c(1610), c(1, (n-18):(n-15))), row.names = FALSE, caption = "One case with significantly outlying scores for vitamins B6.")
```

## Vitamin B2
$~$

```{r vitam_B2}
chunk <- data[, c(1, (n-22):(n-19))]
```

(ref:vitaminsB2) Multivariate outlier analysis for data on vitamin B2. 

```{r vitamins-B2, results='asis', fig.cap="(ref:vitaminsB2)", fig.height=11, fig.width=10, cache=TRUE}
multv_outliers(chunk)
```


```{r vitamins-B2-table, cache=TRUE}
knitr::kable(subset(data, kood %in% c(1139, 1404, 1140, 1610, 1546), c(1, (n-22):(n-19))), row.names = FALSE, caption = "Five cases with significantly outlying scores for vitamin B2.")
```

## Vitamin B12
$~$

```{r vitam_B12}
chunk <- data[, c(1, (n-26):(n-23))]
```

(ref:vitaminsB12) Multivariate outlier analysis for data on vitamin B12. 

```{r vitamins-B12, results='asis', fig.cap="(ref:vitaminsB12)", fig.height=11, fig.width=10, cache=TRUE}
multv_outliers(chunk)
```


```{r vitamins-B12-table, cache=TRUE}
knitr::kable(subset(data, kood %in% c(1054, 1200, 1714, 1706), c(1, (n-26):(n-23))), row.names = FALSE, caption = "Four cases with significantly outlying scores for vitamin B12.")
```

## Vitamin B1
$~$

```{r vitam_B1}
chunk <- data[, c(1, (n-30):(n-27))]
```

(ref:vitaminsB1) Multivariate outlier analysis for data on vitamin B1. 

```{r vitamins-B1, results='asis', fig.cap="(ref:vitaminsB1)", fig.height=11, fig.width=10, cache=TRUE}
multv_outliers(chunk)
```


```{r vitamins-B1-table, cache=TRUE}
knitr::kable(subset(data, kood %in% c(1140, 1610), c(1, (n-30):(n-27))), row.names = FALSE, caption = "Two cases with significantly outlying scores for vitamin B1.")
```

## Vitamin A
$~$

```{r vitam_A}
chunk <- data[, c(1, (n-34):(n-31))]
```

(ref:vitaminsA) Multivariate outlier analysis for data on vitamin A. 

```{r vitamins-A, results='asis', fig.cap="(ref:vitaminsA)", fig.height=11, fig.width=10, cache=TRUE}
multv_outliers(chunk)
```


```{r vitamins-A-table, cache=TRUE}
knitr::kable(subset(data, kood %in% c(1039, 1159, 1530, 1534, 2024, 1706, 1852, 1714, 1200), c(1, (n-34):(n-31))), row.names = FALSE, caption = "Nine cases with significantly outlying scores for vitamin A.")
```

## Vegetables
$~$

```{r veg}
chunk <- data[, c(1, (n-38):(n-35))]
```

(ref:vegetables) Multivariate outlier analysis for data on vegetables. 

```{r veget, results='asis', fig.cap="(ref:vegetables)", fig.height=11, fig.width=10, cache=TRUE}
multv_outliers(chunk)
```


```{r veget-table, cache=TRUE}
knitr::kable(subset(data, kood %in% c(1572, 1830, 1169), c(1, (n-38):(n-35))), row.names = FALSE, caption = "Three cases with significantly outlying scores for vegetables.")
```

## Retinol
$~$
Retinol data has a lot of missing values among 18 years old subjects. 

## Proteins
$~$
Protein data has no cases of putative outliers.

## Poultry
$~$
All values for poultry at age 25 are zeros if poultry values exist for age 33.

```{r poultry}
chunk <- data_poultry <- data[, c(1, (n-59):(n-58))] %>% na.omit()
```

```{r poultry-table, cache=TRUE}
knitr::kable(chunk[1:10, ], row.names = FALSE, caption = "At age 25 in the older cohort all poultry values are 0. (Data of first 10 cases are illustrated).")
```

## Pantothen
$~$
No data for older cohort at 18 years of age. 
Otherwise the data looks OK.

## Other drinks
$~$
Non-alcoholic drinks data has no cases of putative outliers.

## Niacin
$~$
Niacin data has no cases of putative outliers.

## Dairy
$~$

```{r dairy}
chunk <- data[, c(1, (n-79):(n-76))]
```

(ref:dairy-fig) Multivariate outlier analysis for data on dairy. 

```{r dairy-figure, results='asis', fig.cap="(ref:dairy-fig)", fig.height=11, fig.width=10, cache=TRUE}
multv_outliers(chunk)
```


```{r dairy-table, cache=TRUE}
knitr::kable(subset(data, kood %in% c(1176), c(1, (n-79):(n-76))), row.names = FALSE, caption = "One case with significantly outlying scores for dairy.")
```

## Meat
$~$
Meat data has no cases of putative outliers.

## Lipids
$~$
Overall the data looks to be without outliers.

```{r lipids}
chunk <- data[, c(1, (n-87):(n-84))]
```

```{r lipids-table, cache=TRUE}
knitr::kable(subset(data, kood %in% c(1546), c(1, (n-87):(n-84))), row.names = FALSE, caption = "One case with increasingly high scores for lipids.")
```

## LDL
$~$
LDL data has no measurements for 15 years old subjects, otherwise there are no cases of multivariate outliers.

## HOMA
$~$

```{r homa}
chunk <- data[, c(1, (n-99):(n-96))]
```

(ref:homa-fig) Multivariate outlier analysis for data on HOMA index. 

```{r homa-figure, results='asis', fig.cap="(ref:homa-fig)", fig.height=11, fig.width=10, cache=TRUE}
multv_outliers(chunk)
```


```{r homa-table, cache=TRUE}
knitr::kable(subset(data, kood %in% c(1843, 1028, 1589), c(1, (n-99):(n-96))), row.names = FALSE, caption = "Three cases with significantly outlying scores for HOMA index.")
```

## HDL
$~$
HDL data has no cases of putative outliers.

## Fruits and berries
$~$

```{r fruit}
chunk <- data[, c(1, (n-107):(n-104))]
```

(ref:fruit-fig) Multivariate outlier analysis for data on fruits and berries. 

```{r fruit-figure, results='asis', fig.cap="(ref:fruit-fig)", fig.height=11, fig.width=10, cache=TRUE}
multv_outliers(chunk)
```


```{r fruit-table, cache=TRUE}
knitr::kable(subset(data, kood %in% c(1429), c(1, (n-107):(n-104))), row.names = FALSE, caption = "One case with significantly outlying scores for fruits and berries.")
```

## Folate
$~$

```{r folate}
chunk <- data[, c(1, (n-111):(n-108))]
```

(ref:folate-fig) Multivariate outlier analysis for data on folate. 

```{r folate-figure, results='asis', fig.cap="(ref:folate-fig)", fig.height=11, fig.width=10, cache=TRUE}
multv_outliers(chunk)
```


```{r folate-table, cache=TRUE}
knitr::kable(subset(data, kood %in% c(1168, 1860), c(1, (n-111):(n-108))), row.names = FALSE, caption = "Two cases with significantly outlying scores for folate.")
```

## Fish
$~$
Fish data has a lot of 0 values, the multivariate test fails to converge.

## Fats
$~$
Fats data has no cases of putative outliers.

## Eggs
$~$

```{r eggs}
chunk <- data[, c(1, (n-123):(n-120))]
```

(ref:eggs-fig) Multivariate outlier analysis for data on eggs. 

```{r eggs-figure, results='asis', fig.cap="(ref:eggs-fig)", fig.height=11, fig.width=10, cache=TRUE}
multv_outliers(chunk)
```


```{r eggs-table, cache=TRUE}
knitr::kable(subset(data, kood %in% c(1825, 2029), c(1, (n-123):(n-120))), row.names = FALSE, caption = "Two cases with significantly outlying scores for eggs.")
```


## CHL
$~$

```{r chl}
chunk <- data[, c(1, (n-131):(n-128))]
```

(ref:chl-fig) Multivariate outlier analysis for data on CHL. 

```{r chl-figure, results='asis', fig.cap="(ref:chl-fig)", fig.height=11, fig.width=10, cache=TRUE}
multv_outliers(chunk, tolBAC=0.00001, toladj=0.1, tolDDC=0.01)
```


```{r chl-table, cache=TRUE}
knitr::kable(subset(data, kood %in% c(1594, 1597), c(1, (n-131):(n-128))), row.names = FALSE, caption = "Two cases with significantly outlying scores for CHL.")
```

## Cereal
$~$
Cereal data has no cases of putative outliers.

## Carbohydrates
$~$
Carbohydrates data has no cases of putative outliers.

## Biotin
$~$
Biotin data has no measurements for 18 years old subjects.

```{r biotin}
chunk <- data[, c(1, (n-148):(n-147))]
```

(ref:biotin-fig) Multivariate outlier analysis for data on biotin. 

```{r biotin-figure, results='asis', fig.cap="(ref:biotin-fig)", fig.height=11, fig.width=10, cache=TRUE}
multv_outliers(chunk)
```


```{r biotin-table, cache=TRUE}
knitr::kable(subset(data, kood %in% c(1546), c(1, (n-148):(n-147))), row.names = FALSE, caption = "One case with significantly high scores for biotin.")
```

## Alcohol
$~$
Alcohol data has a lot of 0 values, the multivariate test fails to converge.
