---
title             : "ELIKTU univariate analysis of outliers"
shorttitle        : "outliers"
author:
  - name          : "Denis Matrov"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "jazzhole@gmail.com"

affiliation:
  - id            : "1"
    institution   : "Tallinn University"
    
keywords          : ["outliers", "eliktu"]
wordcount         : "X"

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, echo = FALSE, include = FALSE, cache=FALSE}
library("papaja")
library("ggplot2")
library("scales")
library("broom")
library("kableExtra")
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(
  cache = TRUE,
  cache.comments = FALSE,
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  echo = FALSE,
  strip.white = TRUE,
  comment = "#>",
  fig.path = "../figures/",
  results = "asis",
  tidy = "styler",
  dev = c('pdf', 'tiff'),
  dev.args = list(pdf = list(colormodel = 'cmyk'), 
                  tiff = list(compression = 'lzw')),
  dpi = 600,
  fig.width = 7,
  fig.height = 7,
  concordance = TRUE,
  global.par = TRUE
)

par(font.main = 1, cex.main = 1.05)
options(knitr.table.format = "latex")
theme_set(theme_apa(base_size = 15) + theme(legend.position = "bottom"))
```


# Univariate analysis of outliers

```{r data-load, cache=FALSE}
library(here)
library(tidyverse)
library(OutlierDetection)
options(scipen=999)
path <- here("mod_data", "AMIS_toitumine_wide.csv")
data <- read_csv(path)
data <- data[, -grep("_9", colnames(data))]
col_names <- colnames(data)
row_names <- data[, 1, drop = TRUE]
```

## HOMA_15
Subject 1196 has zero-valued HOMA index


## LDL_33
4 values of LDL are 0s

```{r ldl-33, cache=TRUE}
test_data <- data[ ,266, drop = TRUE]
test_data_na <- test_data[!is.na(test_data)]
row_names_na <- row_names
row_names_na <- row_names_na[!is.na(test_data)]
names(test_data_na) <- row_names_na
g1 <- UnivariateOutlierDetection(test_data_na, cutoff = .99, dens = FALSE, depth = TRUE, dist = TRUE, Method = "euclidean", rnames = TRUE)
g1$`Scatter plot` + labs(x = "id", y = "value", title = "")
```

## HDL_33
4 values of HDL are 0s

```{r hdl-33, cache=TRUE}
test_data <- data[ ,265, drop = TRUE]
test_data_na <- test_data[!is.na(test_data)]
row_names_na <- row_names
row_names_na <- row_names_na[!is.na(test_data)]
names(test_data_na) <- row_names_na
g1 <- UnivariateOutlierDetection(test_data_na, cutoff = .99, dens = FALSE, depth = TRUE, dist = TRUE, Method = "euclidean", rnames = TRUE)
g1$`Scatter plot` + labs(x = "id", y = "value", title = "")
```

## CHL_33
4 values of CHL are 0s

```{r chl-33, cache=TRUE}
test_data <- data[ ,264, drop = TRUE]
test_data_na <- test_data[!is.na(test_data)]
row_names_na <- row_names
row_names_na <- row_names_na[!is.na(test_data)]
names(test_data_na) <- row_names_na
g1 <- UnivariateOutlierDetection(test_data_na, cutoff = .99, dens = FALSE, depth = TRUE, dist = TRUE, Method = "euclidean", rnames = TRUE)
g1$`Scatter plot` + labs(x = "id", y = "value", title = "")
```

## HDL_15
Raw values look binned.

```{r hdl-15, cache=TRUE}
test_data <- data[ ,256 , drop = TRUE]
test_data_na <- test_data[!is.na(test_data)]
row_names_na <- row_names
row_names_na <- row_names_na[!is.na(test_data)]
names(test_data_na) <- row_names_na
g1 <- UnivariateOutlierDetection(test_data_na, cutoff = .99, dens = FALSE, depth = TRUE, dist = TRUE, Method = "euclidean", rnames = TRUE)
g1$`Scatter plot` + labs(x = "id", y = "value", title = "")
```

```{r chl-15, cache=TRUE}
test_data <- data[ ,255 , drop = TRUE]
test_data_na <- test_data[!is.na(test_data)]
row_names_na <- row_names
row_names_na <- row_names_na[!is.na(test_data)]
names(test_data_na) <- row_names_na
g1 <- UnivariateOutlierDetection(test_data_na, cutoff = .99, dens = FALSE, depth = TRUE, dist = TRUE, Method = "euclidean", rnames = TRUE)
g1$`Scatter plot` + labs(x = "id", y = "value", title = "")
```

## Cholesterol
One case of total(?) cholesterol had a significantly low value: table \@ref(tab:cholesterol-table)
```{r cholesterol-table, cache=TRUE}
knitr::kable(subset(data, kood %in% c(1385), c(1, 255:257)), row.names = FALSE, caption = "One case with a very small value for cholesterol")
```

## Alcohol_15
This data distribution looks normal to me, I included it for comparison with later ages.

```{r alco-15, cache=TRUE}
test_data <- data[ ,145 , drop = TRUE]
test_data_na <- test_data[!is.na(test_data)]
row_names_na <- row_names
row_names_na <- row_names_na[!is.na(test_data)]
names(test_data_na) <- row_names_na
g1 <- UnivariateOutlierDetection(test_data_na, cutoff = .99, dens = FALSE, depth = TRUE, dist = TRUE, Method = "euclidean", rnames = TRUE)
g1$`Scatter plot` + labs(x = "id", y = "value", title = "")
```

## Alcohol_18
Two values of 3 kg of alcoholic drinks.

```{r alco-18, cache=TRUE}
test_data <- data[ ,182 , drop = TRUE]
test_data_na <- test_data[!is.na(test_data)]
row_names_na <- row_names
row_names_na <- row_names_na[!is.na(test_data)]
names(test_data_na) <- row_names_na
g1 <- UnivariateOutlierDetection(test_data_na, cutoff = .99, dens = FALSE, depth = TRUE, dist = TRUE, Method = "euclidean", rnames = TRUE)
g1$`Scatter plot` + labs(x = "id", y = "value", title = "")
```

## Alcohol_25
One value is of more than 3 kg of alcoholic drinks.

```{r alco-25, cache=TRUE}
test_data <- data[ ,221 , drop = TRUE]
test_data_na <- test_data[!is.na(test_data)]
row_names_na <- row_names
row_names_na <- row_names_na[!is.na(test_data)]
names(test_data_na) <- row_names_na
g1 <- UnivariateOutlierDetection(test_data_na, cutoff = .99, dens = FALSE, depth = TRUE, dist = TRUE, Method = "euclidean", rnames = TRUE)
g1$`Scatter plot` + labs(x = "id", y = "value", title = "")
```

## Alcohol_33
One value at 3 kg and another value at 5 kg of alcoholic drinks. The people who had high values at age 25 are generally no longer outliers this time, the same applies for responses given at age 18. To explore the situation further, I also binned Nitridata results into 3 bins to make them comparable to the questionnaire data about light alcohol. The plot shows that Nutridata results are generally in a good agreement with questionnaire data as there are no cases where a person selected option '1' in the questionnaire and showed high alcohol intake in the diary. The questionnaire seems to be a better indicator of the alcohol consumption, as Nutridata results are more prone to the situational fluctuations. Similar results were observed for ages 18 and 25.

```{r alco-33, cache=TRUE}
test_data <- data[ ,254 , drop = TRUE]
test_data_na <- test_data[!is.na(test_data)]
row_names_na <- row_names
row_names_na <- row_names_na[!is.na(test_data)]
names(test_data_na) <- row_names_na
g1 <- UnivariateOutlierDetection(test_data_na, cutoff = .99, dens = FALSE, depth = TRUE, dist = TRUE, Method = "euclidean", rnames = TRUE)
g1$`Scatter plot` + labs(x = "id", y = "value", title = "")
```

```{r alco_quest-33, cache=TRUE}
data$alko_c_33 <- cut(data$Alcohol_33, 3, labels = c("1", "2", "3"), ordered_result = TRUE)
alko_data_33 <- data %>% select(alko_c_33, lahja_alko_33) %>% filter(!is.na(.))
g <- ggplot(alko_data_33, aes(x = alko_c_33, y = lahja_alko_33)) + 
  geom_jitter(alpha = 0.5, width = 0.16, height = 0.16) + 
  labs(x = "alcohol Nutridata", y = "light alcohol questionnaire", title = "33 years of age")
g
```


## Tea_18
One value is at 3 kg.

```{r tea-18, cache=TRUE}
test_data <- data[ ,181 , drop = TRUE]
test_data_na <- test_data[!is.na(test_data)]
row_names_na <- row_names
row_names_na <- row_names_na[!is.na(test_data)]
names(test_data_na) <- row_names_na
g1 <- UnivariateOutlierDetection(test_data_na, cutoff = .99, dens = FALSE, depth = TRUE, dist = TRUE, Method = "euclidean", rnames = TRUE)
g1$`Scatter plot` + labs(x = "id", y = "value", title = "")
```

## Otherdrinks_33
One value is really very high at 6.8 kg.

```{r otherdrinks-33, cache=TRUE}
test_data <- data[ ,253 , drop = TRUE]
test_data_na <- test_data[!is.na(test_data)]
row_names_na <- row_names
row_names_na <- row_names_na[!is.na(test_data)]
names(test_data_na) <- row_names_na
g1 <- UnivariateOutlierDetection(test_data_na, cutoff = .99, dens = FALSE, depth = TRUE, dist = TRUE, Method = "euclidean", rnames = TRUE)
g1$`Scatter plot` + labs(x = "id", y = "value", title = "")
```

## Eggs_33
One value is very high.

```{r eggs-33, cache=TRUE}
test_data <- data[ ,251 , drop = TRUE]
test_data_na <- test_data[!is.na(test_data)]
row_names_na <- row_names
row_names_na <- row_names_na[!is.na(test_data)]
names(test_data_na) <- row_names_na
g1 <- UnivariateOutlierDetection(test_data_na, cutoff = .99, dens = FALSE, depth = TRUE, dist = TRUE, Method = "euclidean", rnames = TRUE)
g1$`Scatter plot` + labs(x = "id", y = "value", title = "")
```

## Dairy_25
One value is rather high. When binned Nutridata results are plotted against the questionnaire scores, the graph looks reasonable, the number of observations in the top left corner is rather large. 

```{r dairy-25, cache=TRUE}
test_data <- data[ ,214 , drop = TRUE]
test_data_na <- test_data[!is.na(test_data)]
row_names_na <- row_names
row_names_na <- row_names_na[!is.na(test_data)]
names(test_data_na) <- row_names_na
g1 <- UnivariateOutlierDetection(test_data_na, cutoff = .99, dens = FALSE, depth = TRUE, dist = TRUE, Method = "euclidean", rnames = TRUE)
g1$`Scatter plot` + labs(x = "id", y = "value", title = "")
```

```{r dairy_quest-25, cache=TRUE}
data$Milkprod_c_25 <- cut(data$Milkprod_25, 3, labels = c("1", "2", "3"), 
                          ordered_result = TRUE)
dairy_data_25 <- data %>% select(Milkprod_c_25, piimatooted_25) %>% filter(!is.na(.))
ggplot(dairy_data_25, aes(x = Milkprod_c_25, y = piimatooted_25)) + 
  geom_jitter(alpha = 0.5, width = 0.16, height = 0.16) + 
  labs(x = "dairy Nutridata", y = "dairy questionnaire", title = "25 years of age")  
```


## Dairy_33
One value is rather high. The same person was an outlier at age of 25 years.

```{r dairy-33, cache=TRUE}
test_data <- data[ ,247 , drop = TRUE]
test_data_na <- test_data[!is.na(test_data)]
row_names_na <- row_names
row_names_na <- row_names_na[!is.na(test_data)]
names(test_data_na) <- row_names_na
g1 <- UnivariateOutlierDetection(test_data_na, cutoff = .99, dens = FALSE, depth = TRUE, dist = TRUE, Method = "euclidean", rnames = TRUE)
g1$`Scatter plot` + labs(x = "id", y = "value", title = "")
```

## SugSweets_25
One value is very high.

```{r sugsweets-25, cache=TRUE}
test_data <- data[ ,219 , drop = TRUE]
test_data_na <- test_data[!is.na(test_data)]
row_names_na <- row_names
row_names_na <- row_names_na[!is.na(test_data)]
names(test_data_na) <- row_names_na
g1 <- UnivariateOutlierDetection(test_data_na, cutoff = .99, dens = FALSE, depth = TRUE, dist = TRUE, Method = "euclidean", rnames = TRUE)
g1$`Scatter plot` + labs(x = "id", y = "value", title = "")
```

## FruitsBerries_25
One value is very high.

```{r fruitsberries-25, cache=TRUE}
test_data <- data[ ,212 , drop = TRUE]
test_data_na <- test_data[!is.na(test_data)]
row_names_na <- row_names
row_names_na <- row_names_na[!is.na(test_data)]
names(test_data_na) <- row_names_na
g1 <- UnivariateOutlierDetection(test_data_na, cutoff = .99, dens = FALSE, depth = TRUE, dist = TRUE, Method = "euclidean", rnames = TRUE)
g1$`Scatter plot` + labs(x = "id", y = "value", title = "")
```

## Veget_15
Some values are quite high.

```{r veget-15, cache=TRUE}
test_data <- data[ ,133 , drop = TRUE]
test_data_na <- test_data[!is.na(test_data)]
row_names_na <- row_names
row_names_na <- row_names_na[!is.na(test_data)]
names(test_data_na) <- row_names_na
g1 <- UnivariateOutlierDetection(test_data_na, cutoff = .99, dens = FALSE, depth = TRUE, dist = TRUE, Method = "euclidean", rnames = TRUE)
g1$`Scatter plot` + labs(x = "id", y = "value", title = "")
```

## Veget_18
One value is very high.

```{r veget-18, cache=TRUE}
test_data <- data[ ,171 , drop = TRUE]
test_data_na <- test_data[!is.na(test_data)]
row_names_na <- row_names
row_names_na <- row_names_na[!is.na(test_data)]
names(test_data_na) <- row_names_na
g1 <- UnivariateOutlierDetection(test_data_na, cutoff = .99, dens = FALSE, depth = TRUE, dist = TRUE, Method = "euclidean", rnames = TRUE)
g1$`Scatter plot` + labs(x = "id", y = "value", title = "")
```

## Veget_25
One value is very high. I also plotted the binned Nutridata against the questionnaire scores. 

```{r veget-25, cache=TRUE}
test_data <- data[ ,211 , drop = TRUE]
test_data_na <- test_data[!is.na(test_data)]
row_names_na <- row_names
row_names_na <- row_names_na[!is.na(test_data)]
names(test_data_na) <- row_names_na
g1 <- UnivariateOutlierDetection(test_data_na, cutoff = .99, dens = FALSE, depth = TRUE, dist = TRUE, Method = "euclidean", rnames = TRUE)
g1$`Scatter plot` + labs(x = "id", y = "value", title = "")
```

```{r}
data$Veget_c_25 <- cut(data$Veget_25, 3, labels = c("1", "2", "3"), 
                          ordered_result = TRUE)
veget_data_25 <- data %>% select(Veget_c_25, köögivili_25) %>% filter(!is.na(.))
ggplot(veget_data_25, aes(x = Veget_c_25, y = köögivili_25)) + 
  geom_jitter(alpha = 0.5, width = 0.16, height = 0.16) + 
  labs(x = "vegetables Nutridata", y = "vegetables questionnaire", title = "25 years of age")
```


## VitB12_25
One value is very high.

```{r vitb12-25, cache=TRUE}
test_data <- data[ ,204 , drop = TRUE]
test_data_na <- test_data[!is.na(test_data)]
row_names_na <- row_names
row_names_na <- row_names_na[!is.na(test_data)]
names(test_data_na) <- row_names_na
g1 <- UnivariateOutlierDetection(test_data_na, cutoff = .99, dens = FALSE, depth = TRUE, dist = TRUE, Method = "euclidean", rnames = TRUE)
g1$`Scatter plot` + labs(x = "id", y = "value", title = "")
```

## Retinol_18
One value is very high.

```{r retinol-18, cache=TRUE}
test_data <- data[ ,155 , drop = TRUE]
test_data_na <- test_data[!is.na(test_data)]
row_names_na <- row_names
row_names_na <- row_names_na[!is.na(test_data)]
names(test_data_na) <- row_names_na
g1 <- UnivariateOutlierDetection(test_data_na, cutoff = .99, dens = FALSE, depth = TRUE, dist = TRUE, Method = "euclidean", rnames = TRUE)
g1$`Scatter plot` + labs(x = "id", y = "value", title = "")
```

## Retinol_25
One value is very high.

```{r retinol-25, cache=TRUE}
test_data <- data[ ,192 , drop = TRUE]
test_data_na <- test_data[!is.na(test_data)]
row_names_na <- row_names
row_names_na <- row_names_na[!is.na(test_data)]
names(test_data_na) <- row_names_na
g1 <- UnivariateOutlierDetection(test_data_na, cutoff = .99, dens = FALSE, depth = TRUE, dist = TRUE, Method = "euclidean", rnames = TRUE)
g1$`Scatter plot` + labs(x = "id", y = "value", title = "")
```

## Vit_ARE_25
One value is very high.

```{r vitare-25, cache=TRUE}
test_data <- data[ ,191 , drop = TRUE]
test_data_na <- test_data[!is.na(test_data)]
row_names_na <- row_names
row_names_na <- row_names_na[!is.na(test_data)]
names(test_data_na) <- row_names_na
g1 <- UnivariateOutlierDetection(test_data_na, cutoff = .99, dens = FALSE, depth = TRUE, dist = TRUE, Method = "euclidean", rnames = TRUE)
g1$`Scatter plot` + labs(x = "id", y = "value", title = "")
```

## carb_g_18
One value is very high.

```{r carb_g-25, cache=TRUE}
test_data <- data[ ,149 , drop = TRUE]
test_data_na <- test_data[!is.na(test_data)]
row_names_na <- row_names
row_names_na <- row_names_na[!is.na(test_data)]
names(test_data_na) <- row_names_na
g1 <- UnivariateOutlierDetection(test_data_na, cutoff = .99, dens = FALSE, depth = TRUE, dist = TRUE, Method = "euclidean", rnames = TRUE)
g1$`Scatter plot` + labs(x = "id", y = "value", title = "")
```

