---
title             : "ELIKTU univariate analysis of outliers"
shorttitle        : "outliers"
author:
  - name          : "Denis Matrov"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "jazzhole@gmail.com"

affiliation:
  - id            : "1"
    institution   : "Tallinn University"
    
keywords          : ["outliers", "eliktu"]
wordcount         : "X"

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, echo = FALSE, include = FALSE, cache=FALSE}
library("papaja")
library("ggplot2")
library("scales")
library("broom")
library("kableExtra")
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(
  cache = TRUE,
  cache.comments = FALSE,
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  echo = FALSE,
  strip.white = TRUE,
  comment = "#>",
  fig.path = "../figures/",
  results = "asis",
  tidy = "styler",
  dev = c('pdf', 'tiff'),
  dev.args = list(pdf = list(colormodel = 'cmyk'), 
                  tiff = list(compression = 'lzw')),
  dpi = 600,
  fig.width = 7,
  fig.height = 7,
  concordance = TRUE,
  global.par = TRUE
)

par(font.main = 1, cex.main = 1.05)
options(knitr.table.format = "latex")
theme_set(theme_apa(base_size = 15) + theme(legend.position = "bottom"))
```


# Univariate analysis of outliers

```{r data-load, cache=FALSE}
library(here)
library(tidyverse)
library(OutlierDetection)
options(scipen=999)
path <- here("mod_data", "AMIS_toitumine_wide.csv")
data <- read_csv(path)
data <- data[, -grep("_9", colnames(data))]
col_names <- colnames(data)
row_names <- data[, 1, drop = TRUE]
```

## HOMA_15
Subject 1191 has zero-valued HOMA index


## LDL_33
4 values of LDL are 0s

```{r ldl-33, cache=TRUE}
test_data <- data[ ,266, drop = TRUE]
test_data_na <- test_data[!is.na(test_data)]
row_names_na <- row_names
row_names_na <- row_names_na[!is.na(test_data)]
names(test_data_na) <- row_names_na
g1 <- UnivariateOutlierDetection(test_data_na, cutoff = .99, dens = FALSE, depth = TRUE, dist = TRUE, Method = "euclidean", rnames = TRUE)
g1$`Scatter plot` + labs(x = "id", y = "value", title = "")
```

## HDL_33
4 values of HDL are 0s

```{r hdl-33, cache=TRUE}
test_data <- data[ ,265, drop = TRUE]
test_data_na <- test_data[!is.na(test_data)]
row_names_na <- row_names
row_names_na <- row_names_na[!is.na(test_data)]
names(test_data_na) <- row_names_na
g1 <- UnivariateOutlierDetection(test_data_na, cutoff = .99, dens = FALSE, depth = TRUE, dist = TRUE, Method = "euclidean", rnames = TRUE)
g1$`Scatter plot` + labs(x = "id", y = "value", title = "")
```

## CHL_33
4 values of CHL are 0s

```{r chl-33, cache=TRUE}
test_data <- data[ ,264, drop = TRUE]
test_data_na <- test_data[!is.na(test_data)]
row_names_na <- row_names
row_names_na <- row_names_na[!is.na(test_data)]
names(test_data_na) <- row_names_na
g1 <- UnivariateOutlierDetection(test_data_na, cutoff = .99, dens = FALSE, depth = TRUE, dist = TRUE, Method = "euclidean", rnames = TRUE)
g1$`Scatter plot` + labs(x = "id", y = "value", title = "")
```

## HDL_15
Raw values look binned.

```{r hdl-15, cache=TRUE}
test_data <- data[ ,256 , drop = TRUE]
test_data_na <- test_data[!is.na(test_data)]
row_names_na <- row_names
row_names_na <- row_names_na[!is.na(test_data)]
names(test_data_na) <- row_names_na
g1 <- UnivariateOutlierDetection(test_data_na, cutoff = .99, dens = FALSE, depth = TRUE, dist = TRUE, Method = "euclidean", rnames = TRUE)
g1$`Scatter plot` + labs(x = "id", y = "value", title = "")
```

```{r chl-15, cache=TRUE}
test_data <- data[ ,255 , drop = TRUE]
test_data_na <- test_data[!is.na(test_data)]
row_names_na <- row_names
row_names_na <- row_names_na[!is.na(test_data)]
names(test_data_na) <- row_names_na
g1 <- UnivariateOutlierDetection(test_data_na, cutoff = .99, dens = FALSE, depth = TRUE, dist = TRUE, Method = "euclidean", rnames = TRUE)
g1$`Scatter plot` + labs(x = "id", y = "value", title = "")
```

## Cholesterol
One case of total(?) cholesterol had a significantly low value: table \@ref(tab:cholesterol-table)
```{r cholesterol-table, cache=TRUE}
knitr::kable(subset(data, kood %in% c(1385), c(1, 255:257)), row.names = FALSE, caption = "One case with a very small value for cholesterol")
```

## Alcohol_25
One value is of more than 3 kg of alcoholic drinks.

```{r alco-25, cache=TRUE}
test_data <- data[ ,221 , drop = TRUE]
test_data_na <- test_data[!is.na(test_data)]
row_names_na <- row_names
row_names_na <- row_names_na[!is.na(test_data)]
names(test_data_na) <- row_names_na
g1 <- UnivariateOutlierDetection(test_data_na, cutoff = .99, dens = FALSE, depth = TRUE, dist = TRUE, Method = "euclidean", rnames = TRUE)
g1$`Scatter plot` + labs(x = "id", y = "value", title = "")
```

## Alcohol_33
One value at 3 kg and another value at 5 kg of alcoholic drinks. The people who had high values at age 25 are generally no longer outliers this time. 

```{r alco-33, cache=TRUE}
test_data <- data[ ,254 , drop = TRUE]
test_data_na <- test_data[!is.na(test_data)]
row_names_na <- row_names
row_names_na <- row_names_na[!is.na(test_data)]
names(test_data_na) <- row_names_na
g1 <- UnivariateOutlierDetection(test_data_na, cutoff = .99, dens = FALSE, depth = TRUE, dist = TRUE, Method = "euclidean", rnames = TRUE)
g1$`Scatter plot` + labs(x = "id", y = "value", title = "")
```

## Otherdrinks_33
One value is really very high at 6.8 kg.

```{r otherdrinks-33, cache=TRUE}
test_data <- data[ ,253 , drop = TRUE]
test_data_na <- test_data[!is.na(test_data)]
row_names_na <- row_names
row_names_na <- row_names_na[!is.na(test_data)]
names(test_data_na) <- row_names_na
g1 <- UnivariateOutlierDetection(test_data_na, cutoff = .99, dens = FALSE, depth = TRUE, dist = TRUE, Method = "euclidean", rnames = TRUE)
g1$`Scatter plot` + labs(x = "id", y = "value", title = "")
```

## Eggs_33
One value is very high.

```{r eggs-33, cache=TRUE}
test_data <- data[ ,251 , drop = TRUE]
test_data_na <- test_data[!is.na(test_data)]
row_names_na <- row_names
row_names_na <- row_names_na[!is.na(test_data)]
names(test_data_na) <- row_names_na
g1 <- UnivariateOutlierDetection(test_data_na, cutoff = .99, dens = FALSE, depth = TRUE, dist = TRUE, Method = "euclidean", rnames = TRUE)
g1$`Scatter plot` + labs(x = "id", y = "value", title = "")
```

## Dairy_25
One value is rather high.

```{r dairy-25, cache=TRUE}
test_data <- data[ ,214 , drop = TRUE]
test_data_na <- test_data[!is.na(test_data)]
row_names_na <- row_names
row_names_na <- row_names_na[!is.na(test_data)]
names(test_data_na) <- row_names_na
g1 <- UnivariateOutlierDetection(test_data_na, cutoff = .99, dens = FALSE, depth = TRUE, dist = TRUE, Method = "euclidean", rnames = TRUE)
g1$`Scatter plot` + labs(x = "id", y = "value", title = "")
```

## Dairy_33
One value is rather high. The same person was an outlier at age of 25 years.

```{r dairy-33, cache=TRUE}
test_data <- data[ ,247 , drop = TRUE]
test_data_na <- test_data[!is.na(test_data)]
row_names_na <- row_names
row_names_na <- row_names_na[!is.na(test_data)]
names(test_data_na) <- row_names_na
g1 <- UnivariateOutlierDetection(test_data_na, cutoff = .99, dens = FALSE, depth = TRUE, dist = TRUE, Method = "euclidean", rnames = TRUE)
g1$`Scatter plot` + labs(x = "id", y = "value", title = "")
```

## SugSweets_25
One value is very high.

```{r sugsweets-25, cache=TRUE}
test_data <- data[ ,219 , drop = TRUE]
test_data_na <- test_data[!is.na(test_data)]
row_names_na <- row_names
row_names_na <- row_names_na[!is.na(test_data)]
names(test_data_na) <- row_names_na
g1 <- UnivariateOutlierDetection(test_data_na, cutoff = .99, dens = FALSE, depth = TRUE, dist = TRUE, Method = "euclidean", rnames = TRUE)
g1$`Scatter plot` + labs(x = "id", y = "value", title = "")
```
