---
title: "ELIKTU longitudinal mixed-model approach to predicting impulsivity"
author:
- name: Denis Matrov
  affiliation: '1'
  corresponding: yes
  address: Postal address
  email: jazzhole@gmail.com
shorttitle: outliers
output:
  papaja::apa6_pdf: default
  pdf_document: default
keywords:
- outliers
- eliktu
wordcount: X
floatsintext: no
figurelist: no
tablelist: no
footnotelist: no
linenumbers: yes
mask: no
draft: no
documentclass: apa6
classoption: man
affiliation:
- id: '1'
  institution: Tallinn University
---

```{r setup, echo = FALSE, include = FALSE, cache=FALSE}
library("papaja")
library("scales")
library("broom")
library("broom.mixed")
library("here")
library("tidyverse")
library("lme4")
library("kableExtra")
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(
  cache = TRUE,
  cache.comments = FALSE,
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  echo = FALSE,
  strip.white = TRUE,
  comment = "#>",
  fig.path = "../figures/",
  results = "asis",
  tidy = "styler",
  dev = c('pdf', 'tiff'),
  dev.args = list(pdf = list(colormodel = 'cmyk'), 
                  tiff = list(compression = 'lzw')),
  dpi = 600,
  fig.width = 7,
  fig.height = 7,
  concordance = TRUE,
  global.par = TRUE
)

par(font.main = 1, cex.main = 1.05)
options(knitr.table.format = "latex", scipen = 999)
theme_set(theme_apa(base_size = 15) + theme(legend.position = "bottom"))
```

# Does cholesterol predict maladaptive impulsivity?

```{r data-load, cache=TRUE}
path <- here("mod_data", "df_for_lmer_test.csv")
data <- read_csv(path)
data$sugu <- factor(data$sugu, levels = c(1, 2), labels = c("man", "woman"))

# divide data by cohort because cohort 2 does not have CHL_9 scores
# cohort 1 does not have impulsivity scores for age 33, hence I filter data for this age out
kohort_1_df <- data %>% filter(kohort == 1, age != 33) %>% select(-kohort)
kohort_1_df$age_cent <- factor(kohort_1_df$age_cent, levels = c(0, 3, 10),
          labels = c("15 years", "18 years", "25 years"))

# cohort 2 does not have impulsivity scores for age 15, hence I filter data for this age out
kohort_2_df <- data %>% filter(kohort == 2, age != 15) %>% select(-c(kohort, CHL_9))
kohort_2_df$age_cent <- factor(kohort_2_df$age_cent, levels = c(3, 10, 18),
                               labels = c("18 years", "25 years", "33 years"))

```

```{r mimp-chl-cohort1-g, fig.cap="Maladaptive impulsivity scores of cohort 1", cache=TRUE}
g1 <- ggplot(kohort_1_df, aes(x = age, y = Mimp)) + geom_line(alpha = 0.1, aes(group = kood)) +
  stat_summary(fun.data = 'mean_sdl',fun.args = list(mult = 1), geom = 'smooth', se = TRUE,
              colour = "coral", size = 1.8, fill = "palegreen") + 
  scale_x_continuous(breaks = c(15, 18, 25)) +
  labs(y = "Maladaptive impulsivity score",
       title = "Maldaptive impulsivity declined with age in cohort 1") +
  theme(panel.grid.major=element_blank(), panel.grid.minor=element_blank())
g1
```

```{r mimp-chl-cohort2-g, fig.cap="Maladaptive impulsivity scores of cohort 2", cache=TRUE}
g2 <- ggplot(kohort_2_df, aes(x = age, y = Mimp)) + geom_line(alpha = 0.1, aes(group = kood)) +
  stat_summary(fun.data = 'mean_sdl',fun.args = list(mult = 1), geom = 'smooth', se = TRUE,
               colour = "coral", size = 1.8, fill = "palegreen") +
  scale_x_continuous(breaks = c(18, 25, 33)) +
  labs(y = "Maladaptive impulsivity score",
       title = "Maldaptive impulsivity declined with age in cohort 2") +
  theme(panel.grid.major=element_blank(), panel.grid.minor=element_blank())
g2
```

I divided data by cohort because older cohort has no cholesterol scores at age 9 and impulsivity scores at age 15, whereas younger cohort has no impulsivity scores at age 33. Graphical summary of maladaptive impulsivity scores show that the impulsivity declines in the adulthood (Figures \@ref(fig:mimp-chl-cohort1-g) and \@ref(fig:mimp-chl-cohort2-g)). 

Prediction results for cohorts 1 and 2 are illustrated in 2 following tables. Cholesterol did not show a linear effect on maladaptive impulsivity.

```{r lmer_c1_chl}
# model 1 for cohort 1
model_c1_1 <- lmer(Mimp ~ Aimp + age_cent + Wmaxkg + sugu + CHL_9 + CHL +
                            + (1 | kood),
                         REML = TRUE, data = kohort_1_df)
model_c1_summary_1 <- tidy(model_c1_1, ddf = "Kenward-Roger")
kableExtra::kable(model_c1_summary_1, booktabs=TRUE) %>% 
  kable_styling(latex_options="scale_down")
```

```{r lmer_c2_chl}
# model 1 for cohort 1
model_c2_1 <- lmer(Mimp ~ Aimp + age_cent + Wmaxkg + sugu + CHL +
                            + (1 | kood),
                         REML = TRUE, data = kohort_2_df)
model_c2_summary_1 <- tidy(model_c2_1, ddf = "Kenward-Roger")
kableExtra::kable(model_c2_summary_1, booktabs=TRUE) %>% 
  kable_styling(latex_options="scale_down")
```
