---
title             : "ELIKTU: analysis of nutrition and impulsivity"
shorttitle        : "impulsivity"
author:
  - name          : "Denis Matrov"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "jazzhole@gmail.com"

affiliation:
  - id            : "1"
    institution   : "Tallinn University"
    
keywords          : ["food", "eliktu", "impulsivity"]
wordcount         : "X"

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, echo = FALSE, include = FALSE, cache=FALSE}
library("papaja")
library("tidyverse")
library("scales")
library("broom")
library("kableExtra")
library("here")
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(
  cache = TRUE,
  cache.comments = FALSE,
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  error = FALSE,
  echo = FALSE,
  strip.white = TRUE,
  comment = "#>",
  fig.path = "../figures/",
  results = "asis",
  tidy = "styler",
  dev = c('pdf', 'tiff'),
  dev.args = list(pdf = list(colormodel = 'cmyk'), 
                  tiff = list(compression = 'lzw')),
  dpi = 600,
  fig.width = 7,
  fig.height = 7,
  concordance = TRUE,
  global.par = TRUE
)

par(font.main = 1, cex.main = 1.05)
options(knitr.table.format = "latex")
theme_set(theme_apa(base_size = 15) + theme(legend.position = "bottom"))
```

```{r data-prep}
path_a <- here("summary_data", "lmer_results", "aImp_c_both.rds")
path_m <- here("summary_data", "lmer_results", "mImp_c_both.rds")
df_a <- read_rds(path_a)
df_m <- read_rds(path_m)
```

# Predictors of maladaptive impulsivity scores
The missing data was imputed by Blimp software separately for nutritional and impulsivity variables. Data was also divided and imputed separately by cohort. 
The final model was chosen by the lowest AIC fit score. Statistically non-significant variables were included in the final model if their inclusion improved the AIC score. 

```{r table-mImp}
knitr::kable(round(df_m, 3), row.names = TRUE, caption = "Predictors of maladaptive impulsivity score") %>% kable_styling(bootstrap_options = c("striped", "condensed"))
```

# Predictors of adaptive impulsivity scores
The missing data was imputed by Blimp software separately for nutritional and impulsivity variables. Data was also divided and imputed separately by cohort. 
The final model was chosen by the lowest AIC fit score. Statistically non-significant variables were included in the final model if their inclusion improved the AIC score. 

```{r table-aImp}
knitr::kable(round(df_a, 3), row.names = TRUE, caption = "Predictors of adaptive impulsivity score") %>% kable_styling(bootstrap_options = c("striped", "condensed"))
```

